# ISLR Chapter 2 Exercises: Statistical Learning

Answers checked against: https://blog.princehonest.com/stat-learning/

Note: this markdown is not intended to be knit.
```{r}
library(ISLR)
library(tidyverse)
library(MASS)
select <- dplyr::select
```

1a. flexible > inflexible, high n so fit should be more reliable
b. inflexible > flexible, low n might lead to high test MSE
c. flexible > inflexible, flexible methods can follow non-linear fits better
d. inflexible > flexible, to avoid overfitting

7a.
```{r}
c2e7 <- tibble(X1=c(0,2,0,0,-1,1, 0), X2=c(3,0,1,1,0,1, 0), X3=c(0,0,3,2,1,1, 0), 
                      Y=c("R", "R", "R", "G", "G", "R", "test"))
dist(c2e7)
c2e7 <- c2e7 %>% mutate(Euc = 
                          sqrt((c2e7$X1 - c2e7$X1[7])^2 + (c2e7$X2 - c2e7$X2[7])^2 + (c2e7$X3 - c2e7$X3[7])^2))
c2e7
```
b. when K=1, Y=Green, green is closest to Y (obs 5)
c. when K=3, Y=Red, 2x reds and 1x green closest to Y

8a,b.
```{r}
College
```

8cii.
```{r}
pairs(College[,1:10])
```

8ciii.
```{r}
boxplot(College$Private, College$Outstate)
```

8civ.
```{r}
College <- College %>% mutate(Elite = ifelse(Top10perc > 50, "Yes", "No") %>% as.factor())
summary(College)
plot(College$Elite, College$Outstate)
```

9a.
```{r}
auto <- Auto %>% 
  mutate(horsepower = horsepower %>% as.character() %>% as.numeric(),
         cylinders = cylinders %>% as.factor(),
         origin = origin %>% as.factor(),
         year = year %>% as.factor()) %>% 
  drop_na()

auto
```

9b,c
```{r}
get_summary <- function(df, column_pos){
  df %>% 
    gather("predictor", "value", column_pos) %>% 
    group_by(predictor) %>%
    summarise(min = min(value),
              max = max(value),
              mean = mean(value),
              sd = sd(value))
}
get_summary(auto, c(1, 3:6))
```

9d.
```{r}
get_summary(auto[c(1:9, 86:nrow(auto)), ], c(1, 3:6))
```

9e.
```{r}
pairs(auto %>% select(-name))
```

10a,b.
```{r}
pairs(Boston)
```

10g.
```{r}
get_summary(Boston, c(1:length(Boston))) %>% 
  left_join(Boston %>% 
              filter(medv == min(medv), crim > 50) %>% 
              gather("predictor", "value"))
```

10h.
```{r}
Boston %>% mutate(rooms = round(rm)) %>% group_by(rooms) %>% summarise(n = n())
```